---
title: "In-class_EX5"
---

# Get Started

```{r}
#devtools::install_github("LukeCe/spflow")
```

```{r}
pacman::p_load(tmap,sf,spdep,sp,Matrix,spflow,reshape2,knitr,tidyverse)
```

# Data Preparation

## Building the geographical area
MPSZ-2019 shapefile will be import into R environment as a sf tibble data.frame called mpsz.

```{r}
mpsz <- st_read(dsn = "data/geospatial",
                layer = "MPSZ-2019") %>%
  st_transform(crs = 3414) 
```

the code chunk below will be used to import BusStop shapefile into R environment as an sf object called busstop.
```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)

```

our analysis will be focused on planning subzone with bus stop. In view of this, the code chunk below will be used to perform Point-in-Polygon count analysis.
```{r}
mpsz$`BUSSTOP_COUNT`<- lengths(
  st_intersects(
    mpsz, busstop))

```

```{r}


mpsz_busstop <- mpsz %>%
  filter(BUSSTOP_COUNT > 0)
mpsz_busstop

```

## Preparing the Spatial Weights

Code chunk below will be used to compute the three spatial weights at one goal.

```{r}
centroids <- suppressWarnings({
    st_point_on_surface(st_geometry(mpsz_busstop))})

mpsz_nb <- list(
  "by_contiguity" = poly2nb(mpsz_busstop),
  "by_distance" = dnearneigh(centroids, 
                             d1 = 0, d2 = 5000),
  "by_knn" = knn2nb(knearneigh(centroids, 3))
)

```
Code chunks below will be used to plot the spatial weights in mpsz_nb.

```{r}
plot(st_geometry(mpsz))
plot(mpsz_nb$by_contiguity, 
     centroids, 
     add = T, 
     col = rgb(0,0,0,
               alpha=0.5))
title("Contiguity") 

plot(st_geometry(mpsz))
plot(mpsz_nb$by_distance,
     centroids, 
     add = T, 
     col = rgb(0,0,0,
               alpha=0.5)) 
title("Distance") 

plot(st_geometry(mpsz))
plot(mpsz_nb$by_knn, 
     centroids, 
     add = T, 
     col = rgb(0,0,0,
               alpha=0.5))
title("3 Nearest Neighbors") 

```
```{r}
write_rds(mpsz_nb, "data/rds/mpsz_nb.rds")

```

## Preparing The Flow Data

```{r}
odbus6_9 <- read_rds("data/rds/odbus6_9.rds")
```


```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

we are going to append the planning subzone code from busstop_mpsz data.frame onto odbus6_9 data frame.


```{r}
od_data <- left_join(odbus6_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C,
         DESTIN_BS = DESTINATION_PT_CODE)
```

Before continue, it is a good practice for us to check for duplicating records.


```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```


```{r}
od_data <- unique(od_data)
```


```{r}
od_data <- left_join(od_data , busstop_mpsz,
            by = c("DESTIN_BS" = "BUS_STOP_N")) 
```


```{r}
duplicate <- od_data %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```


```{r}
od_data <- unique(od_data)
```


```{r}
od_data <- od_data %>%
  rename(DESTIN_SZ = SUBZONE_C) %>%
  drop_na() %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>%
  summarise(TRIPS = sum(TRIPS))
```

The od_data data.frame should look similar the table below.
```{r}
kable(head(od_data, n = 5))
```


```{r}
write_rds(od_data, "data/rds/od_data.rds")
```

## Computing Distance Matrix
First as.Spatial() will be used to convert mpsz from sf tibble data frame to SpatialPolygonsDataFrame of sp object as shown in the code chunk below.


```{r}
mpsz_sp <- as(mpsz_busstop, "Spatial")
mpsz_sp
```

## Computing the distance matrix
Next, spDists() of sp package will be used to compute the Euclidean distance between the centroids of the planning subzones.
```{r}
DISTANCE <- spDists(mpsz_sp, 
                    longlat = FALSE)
```
Labelling column and row heanders of a distance matrix

```{r}
head(DISTANCE, n=c(10, 10))
```
First, we will create a list sorted according to the the distance matrix by planning sub-zone code.


```{r}
sz_names <- mpsz_busstop$SUBZONE_C
```
Next we will attach SUBZONE_C to row and column for distance matrix matching ahead


```{r}
colnames(DISTANCE) <- paste0(sz_names)
rownames(DISTANCE) <- paste0(sz_names)
```

Pivoting distance value by SUBZONE_C

```{r}
distPair <- melt(DISTANCE) %>%
  rename(DISTANCE = value)
head(distPair, 10)
```

```{r}
distPair <- distPair %>%
  rename(ORIGIN_SZ = Var1,
         DESTIN_SZ = Var2)
```


```{r}
flow_data <- distPair %>%
  left_join (od_data) %>%
  mutate(TRIPS = coalesce(TRIPS, 0))
```


```{r}
kable(head(flow_data, n = 10))
```

```{r}
write_rds(flow_data, "data/rds/mpsz_flow.rds")
```

## Preparing Explanatory Variables
The code chunk below will be used to import pop.csv into R environment and save it as an tibble data.frame object called pop.
```{r}
pop <- read_csv("data/aspatial/pop.csv")
```


```{r}
mpsz_var <- mpsz_busstop %>%
  left_join(pop,
            by = c("PLN_AREA_N" = "PA",
                   "SUBZONE_N" = "SZ")) %>%
  select(1:2, 7:11) %>%
  rename(SZ_NAME = SUBZONE_N,
         SZ_CODE = SUBZONE_C)
```

```{r}
kable(head(mpsz_var[, 1:6], n = 6))
```

Deriving explanatory variables using Point-in-Polygon count

```{r}
schools <- read_rds("data/rds/schools.rds")
```


```{r}
mpsz_var$`SCHOOL_COUNT`<- lengths(
  st_intersects(
    mpsz_var, schools))
```



```{r}
business <- st_read(dsn = "data/geospatial",
                    layer = "Business") %>%
  st_transform(crs = 3414)
```

```{r}
retails <- st_read(dsn = "data/geospatial",
                    layer = "Retails") %>%
  st_transform(crs = 3414)
```


```{r}
finserv <- st_read(dsn = "data/geospatial",
                    layer = "FinServ") %>%
  st_transform(crs = 3414)
```


```{r}
entertn <- st_read(dsn = "data/geospatial",
                    layer = "entertn") %>%
  st_transform(crs = 3414)
```


```{r}
fb <- st_read(dsn = "data/geospatial",
                    layer = "F&B") %>%
  st_transform(crs = 3414)
```

```{r}
lr <- st_read(dsn = "data/geospatial",
                    layer = "Liesure&Recreation") %>%
  st_transform(crs = 3414)
```
Then,we will perform Point-in-Polygon analysis for each of these sf object by using the code chunk below.



```{r}
mpsz_var$`BUSINESS_COUNT`<- lengths(
  st_intersects(
    mpsz_var, business))

mpsz_var$`RETAILS_COUNT`<- lengths(
  st_intersects(
    mpsz_var, retails))

mpsz_var$`FINSERV_COUNT`<- lengths(
  st_intersects(
    mpsz_var, finserv))

mpsz_var$`ENTERTN_COUNT`<- lengths(
  st_intersects(
    mpsz_var, entertn))

mpsz_var$`FB_COUNT`<- lengths(
  st_intersects(
    mpsz_var, fb))

mpsz_var$`LR_COUNT`<- lengths(
  st_intersects(
    mpsz_var, lr))
```


```{r}
glimpse(mpsz_var)

```


```{r}
write_rds(mpsz_var, "data/rds/mpsz_var.rds")
```

# Preparing spflow objects
Let us retrieve by using the code chunk below
```{r}
mpsz_nb <- read_rds("data/rds/mpsz_nb.rds")
mpsz_flow <- read_rds("data/rds/mpsz_flow.rds")
mpsz_var <- read_rds("data/rds/mpsz_var.rds")
```

## Creating spflow_network-class objects
For our model, we choose the contiguity based neighborhood structure.
```{r}
mpsz_net <- spflow_network(
  id_net = "sg",
  node_neighborhood = nb2mat(mpsz_nb$by_contiguity),
  node_data = mpsz_var,
  node_key_column = "SZ_CODE")

mpsz_net
```
## Creating spflow_network-class object
In spflow package, spflow_network_pair()

```{r}
mpsz_net_pairs <- spflow_network_pair(
  id_orig_net = "sg",
  id_dest_net = "sg",
  pair_data = mpsz_flow,
  orig_key_column = "ORIGIN_SZ",
  dest_key_column = "DESTIN_SZ")

mpsz_net_pairs
```

## Creating spflow_network_multi-class object

```{r}
mpsz_multi_net <-  spflow_network_multi(mpsz_net,
                                   mpsz_net_pairs)
mpsz_multi_net
```

Given the information on origins, destinations and OD pairs we can use the spflow_map() method for a simple geographic representation of the largest flows.
```{r}
plot(mpsz$geometry)
spflow_map(
  mpsz_multi_net,
  flow_var = "TRIPS",
  add = TRUE, 
  legend_position = "bottomleft",
  filter_lowest = .999, 
  remove_intra = TRUE,
  cex = 1)
```

## Correlation Analysis
In order to avoid including explanatory variables that are highly correlated, spflow provides two functions:

pair_cor() to create a correlation matrix, and
cor_image() to plot the correlation matrix as a correlogram.
```{r}
cor_formula <- log(1 + TRIPS) ~ 
    BUSSTOP_COUNT +
    AGE7_12 +
    AGE13_24 +
    AGE25_64 +
    SCHOOL_COUNT +
    BUSINESS_COUNT +
    RETAILS_COUNT +
    FINSERV_COUNT +
    P_(log(DISTANCE + 1))

cor_mat <- pair_cor(
  mpsz_multi_net, 
  spflow_formula = cor_formula, 
  add_lags_x = FALSE)

colnames(cor_mat) <- paste0(
  substr(
    colnames(cor_mat),1,3),"...")

cor_image(cor_mat)
```

# Model Calibration
## The base model

```{r}
base_model <- spflow(
  spflow_formula = log(1 + TRIPS) ~ 
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
    BUSINESS_COUNT +
    RETAILS_COUNT +
    FINSERV_COUNT) +
    P_(log(DISTANCE + 1)),
  spflow_networks = mpsz_multi_net)

base_model
```


```{r}
plot(base_model)
```

## Residual diagnostics
In the code chunk below, spflow_moran_plots() is used.


```{r}
old_par <- par(mfrow = c(1, 3), 
               mar = c(2,2,2,2))
spflow_moran_plots(base_model)
```
pair_cor() can be used to inspect the relationship of the residual and the explanatory variables by using the code chunk below

```{r}
corr_residual <- pair_cor(base_model)
colnames(corr_residual) <- substr(colnames(corr_residual),1,3)
cor_image(corr_residual)
```


```{r}
model.df <- as_tibble(base_model@spflow_indicators) %>%
  mutate(FITTED_Y = round(exp(FITTED),0))
```

## Working with model control

```{r}
mpsz_flow1 <- mpsz_flow %>%
  left_join(model.df) %>%
  select(1:4,8) %>%
  mutate(diff = (FITTED_Y-TRIPS))
```


```{r}
spflow_formula <- log(1 + TRIPS) ~ 
    O_(BUSSTOP_COUNT +
         AGE25_64) +
    D_(SCHOOL_COUNT +
    BUSINESS_COUNT +
    RETAILS_COUNT +
    FINSERV_COUNT) +
    P_(log(DISTANCE + 1))

model_control <- spflow_control(
  estimation_method = "mle",
  model = "model_8")

mle_model8 <- spflow(
  spflow_formula,
  spflow_networks = mpsz_multi_net,
  estimation_control = model_control)

mle_model8
```


```{r}
old_par <- par(mfrow = c(1, 3), 
               mar = c(2,2,2,2))
spflow_moran_plots(mle_model8)
```


```{r}
par(old_par)
```
