---
title: "Hands-on Exercise 1:Geospatial Data Wrangling with R"
---

## Overview 

In this Hands-on Exercise,I learn how to import and wrangling geospatial data using appropiate R packages.

## Getting Started

The code chunk below install and load sf and tidyverse packages into R environment.

```{r}
pacman::p_load(sf,tidyverse)

```

## import Geospatial Data

### importing polygon feature data

```{r}
mpsz <- st_read(dsn = "data/geospatial",layer="MP14_SUBZONE_WEB_PL")
```

### **Importing polyline feature data in shapefile form**

```{r}
cyclingpath = st_read(dsn = "data/geospatial", 
                         layer = "CyclingPathGazette")
```

### **Importing GIS data in kml format**

```{r}
preschool = st_read("data/geospatial/preschoolslocation.kml")

```

## **Checking the Content of A Simple Feature Data Frame**

### **Working with *st_geometry()***

```{r}
st_geometry(mpsz)
```

### **Working with glimpse**

```{r}
glimpse(mpsz)
```

### **Working with *head***

```{r}
head(mpsz, n=5)  
```
