---
title: "In_class_Ex1.qmd"
---

## Getting Started

Thge code chunk below load the following packages:

-   tmap: for thematic mapping

-   sf: for geospatial data handling.

-   tidyverse: for non-spatial handling.

```{r}
pacman::p_load(tmap, sf, tidyverse, 
               knitr)
```

### Importing the OD data

Firstly we will import the *Passenger Volumne by origin Destination Bus Stops*

```{r}
odbus <- read_csv("data/aspatial/origin_destination_bus_202308.csv")
```

```{r}
glimpse(odbus)
```

```{r}
odbus$ORIGIN_PT_CODE <- as.factor(odbus$ORIGIN_PT_CODE)
odbus$DESTINATION_PT_CODE <- as.factor(odbus$DESTINATION_PT_CODE) 
```

```{r}
glimpse(odbus)
```

### Extracting the study data

```{r}
origin7_9 <- odbus %>%
  filter(DAY_TYPE == "WEEKDAY") %>%
  filter(TIME_PER_HOUR >= 7 &
           TIME_PER_HOUR <= 9) %>%
  group_by(ORIGIN_PT_CODE) %>%
  summarise(TRIPS = sum(TOTAL_TRIPS))
```

```{r}
kable(head(origin7_9))

```

```{r}
write_rds(origin7_9, "data/rds/origin7_9.rds")
```

```{r}
origin7_9 <- read_rds("data/rds/origin7_9.rds")
```

## **Working with Geospatial Data**

### Importing geospatial data

```{r}
busstop <- st_read(dsn = "data/geospatial",
                   layer = "BusStop") %>%
  st_transform(crs = 3414)
```

```{r}
glimpse(busstop)

```

```{r}
mpsz <- st_read(dsn="data/geospatial",layer="MPSZ-2019") %>%
  st_transform(crs=3414)
```

```{r}
glimpse(mpsz)

```

## **Geospatial data wrangling**

### **Combining Busstop and mpsz**

```{r}
busstop_mpsz <- st_intersection(busstop, mpsz) %>%
  select(BUS_STOP_N, SUBZONE_C) %>%
  st_drop_geometry()
```

```{r}
write_rds(busstop_mpsz, "data/rds/busstop_mpsz.csv")  

```

```{r}
origin_SZ <- left_join(origin7_9 , busstop_mpsz,
            by = c("ORIGIN_PT_CODE" = "BUS_STOP_N")) %>%
  rename(ORIGIN_BS = ORIGIN_PT_CODE,
         ORIGIN_SZ = SUBZONE_C) %>%
  group_by(ORIGIN_SZ) %>%
  summarise(TOT_TRIPS = sum(TRIPS))
```

```{r}
duplicate <- origin_SZ %>%
  group_by_all() %>%
  filter(n()>1) %>%
  ungroup()
```

```{r}
origin_data <- unique(origin_SZ)
```

```{r}
origintrip_SZ <- left_join(mpsz, 
                           origin_SZ,
                           by = c("SUBZONE_C" = "ORIGIN_SZ"))
```

## Choropleth Visualisation

```{r}
tm_shape(origintrip_SZ)+
  tm_fill("TOT_TRIPS", 
          style = "quantile", 
          palette = "Blues",
          title = "Passenger trips") +
  tm_layout(main.title = "Passenger trips generated at planning sub-zone level",
            main.title.position = "center",
            main.title.size = 1.2,
            legend.height = 0.45, 
            legend.width = 0.35,
            frame = TRUE) +
  tm_borders(alpha = 0.5) +
  tm_compass(type="8star", size = 2) +
  tm_scale_bar() +
  tm_grid(alpha =0.2) +
  tm_credits("Source: Planning Sub-zone boundary from URA\n and Passenger trips data from LTA", 
             position = c("left", "bottom"))
```
