---
title: "In-class_EX3"
---

# **Calibrating Spatial Interaction Models with R**

## **Overview**

## **The Case Study and Data**

## **Getting Started**

-   sf for importing, integrating, processing and transforming geospatial data.

-   tidyverse for importing, integrating, wrangling and visualising data.

-   tmap for creating thematic maps.

```{r}
pacman::p_load(tmap, sf, sp, DT, stplanr,
               performance, reshape2,
               ggpubr, units, tidyverse)
```

## **16.5 Computing Distance Matrix**

First, let us import *mpsz.rds* into R environemnt by using the code chunk below.

```{r}
mpsz <- read_rds("data/data/rds/mpsz.rds")
mpsz
```

### **16.5.1 Converting from sf data.table to SpatialPolygonsDataFrame**

First [`as.Spatial()`](https://r-spatial.github.io/sf/reference/coerce-methods.html) will be used to convert *mpsz* from sf tibble data frame to SpatialPolygonsDataFrame of sp object as shown in the code chunk below.

```{r}
mpsz_sp <- as(mpsz, "Spatial")
mpsz_sp
```

### **16.5.2 Computing the distance matrix**

Next, [`spDists()`](https://www.rdocumentation.org/packages/sp/versions/2.1-1/topics/spDistsN1) of sp package will be used to compute the Euclidean distance between the centroids of the planning subzones.

```{r}
dist <- spDists(mpsz_sp, 
                longlat = FALSE)
head(dist, n=c(10, 10))
```

### **16.5.3 Labelling column and row heanders of a distance matrix**

First, we will create a list sorted according to the the distance matrix by planning sub-zone code.

```{r}
sz_names <- mpsz$SUBZONE_C
```

Next we will attach `SUBZONE_C` to row and column for distance matrix matching ahead

```{r}
colnames(dist) <- paste0(sz_names)
rownames(dist) <- paste0(sz_names)
```

### **16.5.4 Pivoting distance value by SUBZONE_C**

Next, we will pivot the distance matrix into a long table by using the row and column subzone codes as show in the code chunk below.

```{r}
distPair <- melt(dist) %>%
  rename(dist = value)
head(distPair, 10)
```

### **16.5.5 Updating intra-zonal distances**

First, we will select and find out the minimum value of the distance by using `summary()`.

```{r}
distPair %>%
  filter(dist > 0) %>%
  summary()
```

Next, a constant distance value of 50m is added into intra-zones distance.

```{r}
distPair$dist <- ifelse(distPair$dist == 0,
                        50, distPair$dist)
```

The code chunk below will be used to check the result data.frame.

```{r}
distPair %>%
  summary()
```

The code chunk below is used to rename the origin and destination fields.

```{r}
distPair <- distPair %>%
  rename(orig = Var1,
         dest = Var2)
```

Lastly, the code chunk below is used to save the dataframe for future use.

```{r}
write_rds(distPair, "data/data/rds/distPair.rds") 
```

## **16.6 Preparing flow data**

The code chunk below is used import *od_data* save in Chapter 15 into R environment.

```{r}
od_data <- read_rds("data/data/rds/od_data.rds")
```

Next, we will compute the total passenger trip between and within planning subzones by using the code chunk below. The output is all *flow_data*.

```{r}
flow_data <- od_data %>%
  group_by(ORIGIN_SZ, DESTIN_SZ) %>% 
  summarize(TRIPS = sum(MORNING_PEAK)) 
```

Use the code chunk below to display flow_data dataframe.

```{r}
head(flow_data, 10)
```

### **16.6.1 Separating intra-flow from passenger volume df**

Code chunk below is used to add three new fields in `flow_data` dataframe.

```{r}
flow_data$FlowNoIntra <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0, flow_data$TRIPS)
flow_data$offset <- ifelse(
  flow_data$ORIGIN_SZ == flow_data$DESTIN_SZ, 
  0.000001, 1)
```

### **16.6.2 Combining passenger volume data with distance value**

Before we can join *flow_data* and *distPair*, we need to convert data value type of *ORIGIN_SZ* and *DESTIN_SZ* fields of flow_data dataframe into factor data type.

```{r}
flow_data$ORIGIN_SZ <- as.factor(flow_data$ORIGIN_SZ)
flow_data$DESTIN_SZ <- as.factor(flow_data$DESTIN_SZ)
```

Now, `left_join()` of **dplyr** will be used to *flow_data* dataframe and *distPair* dataframe. The output is called *flow_data1*.

```{r}
flow_data1 <- flow_data %>%
  left_join (distPair,
             by = c("ORIGIN_SZ" = "orig",
                    "DESTIN_SZ" = "dest"))
```

## **16.7 Preparing Origin and Destination Attributes**

### **16.7.1 Importing population data**

```{r}
pop <- read_csv("data/data/aspatial/pop.csv")
```

### **16.7.2 Geospatial data wrangling**

```{r}
pop <- pop %>%
  left_join(mpsz,
            by = c("PA" = "PLN_AREA_N",
                   "SZ" = "SUBZONE_N")) %>%
  select(1:6) %>%
  rename(SZ_NAME = SZ,
         SZ = SUBZONE_C)
```

### **16.7.3 Preparing origin attribute**

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(ORIGIN_SZ = "SZ")) %>%
  rename(ORIGIN_AGE7_12 = AGE7_12,
         ORIGIN_AGE13_24 = AGE13_24,
         ORIGIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

### **16.7.4 Preparing destination attribute**

```{r}
flow_data1 <- flow_data1 %>%
  left_join(pop,
            by = c(DESTIN_SZ = "SZ")) %>%
  rename(DESTIN_AGE7_12 = AGE7_12,
         DESTIN_AGE13_24 = AGE13_24,
         DESTIN_AGE25_64 = AGE25_64) %>%
  select(-c(PA, SZ_NAME))
```

We will called the output data file *SIM_data*. it is in rds data file format.

```{r}
write_rds(flow_data1, "data/data/rds/SIM_data")

```

## **16.8 Calibrating Spatial Interaction Models**

### **16.8.1 Importing the modelling data**

Firstly, let us import the modelling data by using the code chunk below.

```{r}
SIM_data <- read_rds("data/data/rds/SIM_data.rds")
```

### **16.8.2 Visualising the dependent variable**

Firstly, let us plot the distribution of the dependent variable (i.e. TRIPS) by using histogram method by using the code chunk below.

```{r}
ggplot(data = SIM_data,
       aes(x = TRIPS)) +
  geom_histogram()
```

Next, let us visualise the relation between the dependent variable and one of the key independent variable in Spatial Interaction Model, namely distance.

```{r}
ggplot(data = SIM_data,
       aes(x = dist,
           y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)
```

On the other hand, if we plot the scatter plot by using the log transformed version of both variables, we can see that their relationship is more resemble linear relationship.

```{r}
ggplot(data = SIM_data,
       aes(x = log(dist),
           y = log(TRIPS))) +
  geom_point() +
  geom_smooth(method = lm)
```

### **16.8.3 Checking for variables with zero values**

```{r}
summary(SIM_data)

```

In view of this, code chunk below will be used to replace zero values to 0.99.

```{r}
SIM_data$DESTIN_AGE7_12 <- ifelse(
  SIM_data$DESTIN_AGE7_12 == 0,
  0.99, SIM_data$DESTIN_AGE7_12)
SIM_data$DESTIN_AGE13_24 <- ifelse(
  SIM_data$DESTIN_AGE13_24 == 0,
  0.99, SIM_data$DESTIN_AGE13_24)
SIM_data$DESTIN_AGE25_64 <- ifelse(
  SIM_data$DESTIN_AGE25_64 == 0,
  0.99, SIM_data$DESTIN_AGE25_64)
SIM_data$ORIGIN_AGE7_12 <- ifelse(
  SIM_data$ORIGIN_AGE7_12 == 0,
  0.99, SIM_data$ORIGIN_AGE7_12)
SIM_data$ORIGIN_AGE13_24 <- ifelse(
  SIM_data$ORIGIN_AGE13_24 == 0,
  0.99, SIM_data$ORIGIN_AGE13_24)
SIM_data$ORIGIN_AGE25_64 <- ifelse(
  SIM_data$ORIGIN_AGE25_64 == 0,
  0.99, SIM_data$ORIGIN_AGE25_64)
```

```{r}
summary(SIM_data)

```

### **16.8.4 Unconstrained Spatial Interaction Model**

The code chunk used to calibrate to model is shown below:

```{r}
uncSIM <- glm(formula = TRIPS ~ 
                log(ORIGIN_AGE25_64) + 
                log(DESTIN_AGE25_64) +
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
uncSIM
```

### **16.8.5 R-squared function**

In order to measure how much variation of the trips can be accounted by the model we will write a function to calculate R-Squared value as shown below.

```{r}
CalcRSquared <- function(observed,estimated){
  r <- cor(observed,estimated)
  R2 <- r^2
  R2
}
```

Next, we will compute the R-squared of the unconstrained SIM by using the code chunk below.

```{r}
CalcRSquared(uncSIM$data$TRIPS, uncSIM$fitted.values)
```

```{r}
r2_mcfadden(uncSIM)
```

### **16.8.6 Origin (Production) constrained SIM**

In this section, we will fit an origin constrained SIM by using the code3 chunk below.

The general formula of Origin Constrained Spatial Interaction Model.

```{r}
orcSIM <- glm(formula = TRIPS ~ 
                 ORIGIN_SZ +
                 log(DESTIN_AGE25_64) +
                 log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(orcSIM)
```

We can examine how the constraints hold for destinations this time.

```{r}
CalcRSquared(orcSIM$data$TRIPS, orcSIM$fitted.values)

```

### **16.8.7 Destination constrained**

In this section, we will fit a destination constrained SIM by using the code chunk below.

The general formula of Destination Constrained Spatial Interaction Model.

```{r}
decSIM <- glm(formula = TRIPS ~ 
                DESTIN_SZ + 
                log(ORIGIN_AGE25_64) + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(decSIM)
```

We can examine how the constraints hold for destinations this time.

```{r}
CalcRSquared(decSIM$data$TRIPS, decSIM$fitted.values)
```

### **16.8.8 Doubly constrained**

In this section, we will fit a doubly constrained SIM by using the code chunk below.

The general formula of Doubly Constrained Spatial Interaction Model.

```{r}
dbcSIM <- glm(formula = TRIPS ~ 
                ORIGIN_SZ + 
                DESTIN_SZ + 
                log(dist),
              family = poisson(link = "log"),
              data = SIM_data,
              na.action = na.exclude)
summary(dbcSIM)
```

We can examine how the constraints hold for destinations this time.

```{r}
CalcRSquared(dbcSIM$data$TRIPS, dbcSIM$fitted.values)
```

### **16.8.9 Model comparison**

First of all, let us create a list called *model_list* by using the code chun below.

```{r}
model_list <- list(unconstrained=uncSIM,
                   originConstrained=orcSIM,
                   destinationConstrained=decSIM,
                   doublyConstrained=dbcSIM)
```

Next, we will compute the RMSE of all the models in *model_list* file by using the code chunk below.

```{r}
compare_performance(model_list,
                    metrics = "RMSE")
```

### **16.8.10 Visualising fitted**

In this section, you will learn how to visualise the observed values and the fitted values.

Firstly we will extract the fitted values from each model by using the code chunk below.

```{r}
df <- as.data.frame(uncSIM$fitted.values) %>%
  round(digits = 0)
```

Next, we will join the values to *SIM_data* data frame.

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(uncTRIPS = "uncSIM$fitted.values")
```

Repeat the same step by for Origin Constrained SIM (i.e. orcSIM)

```{r}
df <- as.data.frame(orcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(orcTRIPS = "orcSIM$fitted.values")
```

Repeat the same step by for Destination Constrained SIM (i.e. decSIM)

```{r}
df <- as.data.frame(decSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(decTRIPS = "decSIM$fitted.values")
```

Repeat the same step by for Doubly Constrained SIM (i.e. dbcSIM)

```{r}
df <- as.data.frame(dbcSIM$fitted.values) %>%
  round(digits = 0)
```

```{r}
SIM_data <- SIM_data %>%
  cbind(df) %>%
  rename(dbcTRIPS = "dbcSIM$fitted.values")
```

```{r}
unc_p <- ggplot(data = SIM_data,
                aes(x = uncTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

orc_p <- ggplot(data = SIM_data,
                aes(x = orcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dec_p <- ggplot(data = SIM_data,
                aes(x = decTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

dbc_p <- ggplot(data = SIM_data,
                aes(x = dbcTRIPS,
                    y = TRIPS)) +
  geom_point() +
  geom_smooth(method = lm)

ggarrange(unc_p, orc_p, dec_p, dbc_p,
          ncol = 2,
          nrow = 2)
```
